Two NIRCam channels are Better than One: How JWST Can Do More Science with NIRCam's Short-Wavelength Dispersed Hartmann Sensor

### 2016-09-28 ms.tex
re-arranged placement of the resolution and throughput figures and updated some clarifying text.

\subsection{DHS properties}
The DHS expands the wavelength range that NIRCam can achieve on a single observation by simultaneously obtaining spectroscopy with NIRCam's short wavelength channel and long wavelength channel.
The maximum wavelength range for the F150W2 filter in the short wavelength channel is fixed at 1.01 to 2.02$\mu$m with the region from 2.02-2.35$\mu$m contaminated by second order overlap.
The DHS are slitless, so the resolution is set by the pupil sizes of the individual DHS elements, which are approximately the same width in the dispersion direction.
Figure~\ref{fig:DHSRes} shows the resolution as a function of wavelength, estimated from a narrowly peaked input spectrum and the fits to the spatial width along the spectrum.
There is a spatial gap between short wavelength detectors which corresponds to about \DHSgap$\mu$m in wavelength.
The center of this gap depends on source position.
The DHS can be used in tandem with either the F322W2 or F444W filter for the long wavelength grism observations.
The properties of the DHS as compared to the grisms are listed in Table~\ref{tab:DHSgprop}.
If the region from 2-2.23$\mu$m is highly desirable such as for measuring the CH$_4$ feature shown in Figure~\ref{fig:HATp12spec}, the DHS can be used with a F200W filter (1.75 to 2.23$\mu$m half maximum) to partially fill in the gap between the short to long wavelength channels, though the grism transmission is lower above 2.0$\mu$m, shown in Figure~\ref{fig:DHSthrough}.

\begin{table*}[!t]
\centering
\caption{Simultaneous DHS and Grism Observations (Module A)}\label{tab:DHSgprop}
\begin{tabular}{lcc}
\hline \hline
 & DHS & Grism \\
\hline \hline		
Resolution: &  \DHSres & 1200 - 1550 \\
Dispersion: & 0.29 nm / px & 1.0 nm / px \\
Wavelengths: & 1.05 - 2.00$\mu$m (F150W2)\footnote{For a field point matched to the F322W2 filter and long wavelength grism.} & 2.44 - 4.02$\mu$m (F322W2)\\
			 & 1.75 - 2.23$\mu$m (F200W)\footnote{For a field point matched to the F322W2 filter and long wavelength grism.} & 2.44 - 4.02$\mu$m (F322W2)\\
			& 	1.01 - 1.94$\mu$m (F150W2) \footnote{For a field point matched to the F444W filter and long wavelength grism. There is a \DHSgap$\mu$m gap between detectors that depends on field position.}	& 3.88 - 4.98$\mu$m (F444W) \\
Plate Scale: & 32 mas / px  &  65 mas / px\\
Undeviated Wavelength: & 1.36$\mu$m & 3.94$\mu$m   \\
\hline
\end{tabular}
%%%
%\caption{Properties of DHS and simultaneous long-wavelength grism spectra. The parameters are for the A-side grism in the direction aligned with detector rows (R).}\label{tab:DHSgprop}
%\vspace{0.1in}
\end{table*}

\begin{figure}[!b]
\centering
\includegraphics[width=1.0\columnwidth]{dhs_res.pdf}
\caption{Resolution as a function of wavelength for the DHS.}\label{fig:DHSRes}
\end{figure}

\begin{figure}[!b]
\centering
\includegraphics[width=1.0\columnwidth]{NIRCam+OTE_SWA_DHS_and_F150W2_efficiency.pdf}
\caption{Throughput as a function of wavelength for the DHS grism, F150W2 filter and F200W filter that includes the telescope and instrument transmission. The total throughput is the multiple of the filter curve, ``DHS'' grism and the pupil area fraction from Table~\ref{tab:pupfrac}..}\label{fig:DHSthrough}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=0.85\textwidth]{grism_pos_for_max_wcoverage.pdf}
\caption{DHS wavelength coverage for two field positions that will permit grism spectroscopy simultaneously with the F322W2 and F444W filters in the long wavelength channel for the same input source. The V2/V3 axes show the JWST observatory axes and the detectors (A1 - A5) are shown with their (0,0) corners as black boxes. The circles plotted for each grism are the shortest usable wavelength, undispersed wavelength and longest usable wavelength.}\label{fig:DHSwaveWGrism}
\end{figure*}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=1.0\columnwidth]{dhs_pupil_labels_modA.pdf}
\caption{DHS pupil coverage by each of the 10 DHS prisms, with the fractional throughput given in Table~\ref{tab:pupfrac}. 
The throughput is modulated by the size of the DHS grism as well as whether it coincides with a secondary mirror support (inverted "Y" shaped feature).
The pupil wheel is clocked to make the DHS spectra straight along rows.
This moves the DHS elements off-center from JWST's pupil and thus decreases the throughput for elements 1 and 6, but the overall combined throughput for all 10 elements is only reduced from 32\% to 29\%.}\label{fig:DHSvsPupilOverlay}
\end{figure}

The throughput for DHS spectroscopy is the product of the DHS grating efficiency, telescope and instrument throughput for a given filter (both shown in Figure~\ref{fig:DHSthrough}) and the pupil coverage by the DHS elements listed in Table~\ref{tab:pupfrac}.
A fractional pupil 1.00 would use the entire pupil and thus take advantage of the maximum JWST collecting area.
For a sub-array size encompassing only 1 DHS spectrum (such as a 2048 $\times$ 64 subarray), the largest unobscured DHS grisms should be used - element number 7 in Figure~\ref{fig:DHSvsPupilOverlay}.
Only the long-wavelength grisms if Module A are considered, because those in Module B lack an anti-reflective coating.
%We discuss only module A (of the two redundant modules A and B) because the long wavelength grisms are anti-reflective coated for module A.

@@

\subsection{Readout Considerations}\label{sec:readout}
JWST's NIRCam instrument has many different possible readout patterns per pixel and array that all factor in to the SNRs and saturation limits for an observation.
We adopt the following naming convention for the readout: an observation may consist of one of more exposures which are combined into a single \texttt{FITS} file per exposure.
Within an exposure, there are one or more integrations, which consist of a reset frame followed by one or more groups of non-destructive reads up a ramp.
This is illustrated in Figure~\ref{fig:readout} for three example readout modes.

We consider a simplified example of a time series using a standard correlated double sampling (CDS) illustrated in Figure~\ref{fig:readout} and assume a round frame time of 10.0 seconds, close to the actual 10.74 seconds full frame read time.
The raw data delivered to observers for use in a pipeline would consist of a single FITS file containing a data cube.
If the data is in full frame mode, cube would have dimensions of 2048 $\times$ 2048 $\times$ 8, which can be thought of 8 image planes.
In this example there are 4 integrations consisting of a reset followed by two reads, which are saved as two image planes per integration.	
For simplicity in this example, we ignore many of the corrections that are applied in a pipeline such as non-linearity correction, reference pixel correction, intrapixel capacitance, flat fielding etc.
Under these assumptions, the flux (in DN/s) in any integration is the difference between the two reads divided by the frame time.
This reduces the 8 image planes, to 4 flux planes, one per integration.
Spectral extraction on the flux planes \citep[e.g.][]{horne1986optimalE} produces 1 flux per wavelength bin.
The time series for a given wavelength would consist of the 4 fluxes with 4 time coordinates centered at 15 sec, 45 sec, 75 sec and 105 sec if we use the mid-points of the two reads.

@@

There are several variations of readout patterns available to NIRCam with the Astronomers Proposal Tool (APT)\footnote{http://www.stsci.edu/hst/proposing/apt}.
For exoplanets time series observations, the readout patterns most commonly used will likely be ``RAPID'', meaning every sample is saved up the ramp individually and ``BRIGHT1'' where a single frame is dropped between reads.
These modes give high time resolution and allow one to detect and flag cosmic ray hits without having to discard large amounts of data.
The number of groups per integration is a variable set in the APT.
When the number of groups is 1 (``1~Group Readout''), there is no ability to subtract out the bias frame so it relies on the bias level being constant through the whole time series.
When the number of groups is 2, the slope can be calculated from the difference between the frames divided by the frame time.
It is the same as the CDS mode described above when using RAPID readout.
For a higher number groups, a line is fit to the the ramp to calculate a slope image for every integration.
In exoplanet time series, the high signal to noise ratios required generally mean that bright sources will be observed so these will generally not benefit from skipped frames or co-added frames of other JWST readout modes, which come at the expense of lower time resolution, reduced ability for cosmic ray rejection and fainter saturation limits.
For more general discussion of HAWAII-2RG detector readout, we refer to \citet{rauscher2007detectors} and for specific readout patterns we refer to \citet{roberto2009optimalR1} and \citet{roberto2010optimalR2}.

@@

\begin{figure}[!h]
\includegraphics[width=1.0\columnwidth]{brightness_scape.pdf}
\caption{Saturation limits of the DHS mode (blue line), and long-wavelength grism's F322W2 filter (green line) and F444W filter (dark red line) as a function of stellar effective temperature.
The magnitudes are Vega $K_S$ magnitudes where 80\% of the well depth is filled in the standard CDS readout mode (described in Figure~\ref{fig:readout}) for a 2048 $\times$ 64 pixel sub-array using 4 amplifiers.
The blue points are known transiting systems from \texttt{exoplanets.org} and the green points are simulated TESS planets \citep{sullivan2015tess}.}\label{fig:saturationLim}
\end{figure}

@@

The posterior distributions for HD 209458b in Figure~\ref{fig:corner209} show improved constraints on the temperature, cloud pressure level and the abundances of H$_2$O, CO and CO$_2$ when the DHS is used with the long wavelength grisms (Case~\ref{it:DHSgr}) as compared to the long wavelength grisms alone (Case~\ref{it:grAlone}).
These improvements come about, in part, because the 1.0$\mu$m to 2.0$\mu$m wavelength region probes to deeper pressure layers ($\sim$ 400 mbar) to exclude clouds and better constrain the scale height.

>>>>

\subsection{DHS properties}

\begin{figure}
\centering
\includegraphics[width=1.0\columnwidth]{dhs_res.pdf}
\caption{Resolution as a function of wavelength for the DHS.}\label{fig:DHSRes}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1.0\columnwidth]{NIRCam+OTE_SWA_DHS_and_F150W2_efficiency.pdf}
\caption{Throughput as a function of wavelength for the DHS grism, F150W2 filter and F200W filter that includes the telescope and instrument transmission. The total throughput is the multiple of the filter curve, ``DHS'' grism and the pupil area fraction from Table~\ref{tab:pupfrac}..}\label{fig:DHSthrough}
\end{figure}

The DHS expands the wavelength range that NIRCam can achieve on a single observation by simultaneously obtaining spectroscopy with NIRCam's short wavelength channel and long wavelength channel.
The maximum wavelength range for the F150W2 filter in the short wavelength channel is fixed at 1.01 to 2.02$\mu$m with the region from 2.02-2.35$\mu$m contaminated by second order overlap.
The DHS are slitless, so the resolution is set by the pupil sizes of the individual DHS elements, which are approximately the same width as each other in the dispersion direction.
Figure~\ref{fig:DHSRes} shows the resolution as a function of wavelength, estimated from a narrowly peaked input spectrum and the fits to the spatial width along the spectrum.
There is a spatial gap between short wavelength detectors which corresponds to about \DHSgap$\mu$m in wavelength.
The center of this gap depends on source position.
The DHS can be used in tandem with either the F322W2 or F444W filter for the long wavelength grism observations.
The properties of the DHS as compared to the grisms are listed in Table~\ref{tab:DHSgprop}.
If the region from 2-2.23$\mu$m is highly desirable such as for measuring the CH$_4$ feature shown in Figure~\ref{fig:HATp12spec}, the DHS can be used with a F200W filter (1.75 to 2.23$\mu$m half maximum) to partially fill in the gap between the DHS+F150W2 spectrum and Grism+F322W2 spectra, though the grism transmission is lower above 2.0$\mu$m, shown in Figure~\ref{fig:DHSthrough}.

\begin{table*}
\centering
\caption{Simultaneous DHS and Grism Properties (Module A)}\label{tab:DHSgprop}
\begin{tabular}{lcc}
\hline \hline
 & DHS & Grism \\
\hline \hline		
Resolution: &  \DHSres & 1200 - 1550 \\
Dispersion: & 0.29 nm / px & 1.0 nm / px \\
Wavelengths: & 1.05 - 2.00$\mu$m (F150W2)\footnote{For a field point matched to the F322W2 filter and long wavelength grism.} & 2.44 - 4.02$\mu$m (F322W2)\\
			 & 1.75 - 2.23$\mu$m (F200W)\footnote{For a field point matched to the F322W2 filter and long wavelength grism.} & 2.44 - 4.02$\mu$m (F322W2)\\
			& 	1.01 - 1.94$\mu$m (F150W2) \footnote{For a field point matched to the F444W filter and long wavelength grism. There is a \DHSgap$\mu$m gap between detectors that depends on field position.}	& 3.88 - 4.98$\mu$m (F444W) \\
Plate Scale: & 32 mas / px  &  65 mas / px\\
Undeviated Wavelength: & 1.36$\mu$m & 3.94$\mu$m   \\
\hline
\end{tabular}
%%%
%\caption{Properties of DHS and simultaneous long-wavelength grism spectra. The parameters are for the A-side grism in the direction aligned with detector rows (R).}\label{tab:DHSgprop}
%\vspace{0.1in}
\end{table*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.85\textwidth]{grism_pos_for_max_wcoverage.pdf}
\caption{DHS wavelength coverage for two field positions that will permit grism spectroscopy simultaneously with the F322W2 and F444W filters in the long wavelength channel for the same input source. The V2/V3 axes show the JWST observatory axes and the detectors (A1 - A5) are shown with their (0,0) corners as black boxes. The circles plotted for each grism are the shortest usable wavelength, undispersed wavelength and longest usable wavelength.}\label{fig:DHSwaveWGrism}
\end{figure*}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=1.0\columnwidth]{dhs_pupil_labels_modA.pdf}
\caption{DHS pupil coverage by each of the 10 DHS grisms, with the fractional throughput given in Table~\ref{tab:pupfrac}. 
The throughput is modulated by the size of the DHS grism as well as whether it coincides with a secondary mirror support (inverted ``Y'' shaped feature).
The pupil wheel is clocked to make the DHS spectra straight along rows.
This moves the DHS elements off-center from JWST's pupil and thus decreases the throughput for elements 1 and 6, but the overall combined throughput for all 10 elements is only reduced from 32\% to 29\%.}\label{fig:DHSvsPupilOverlay}
\end{figure}

The throughput for DHS spectroscopy is the product of the DHS grating efficiency, telescope and instrument throughput for a given filter (both shown in Figure~\ref{fig:DHSthrough}) and the pupil coverage by the DHS elements listed in Table~\ref{tab:pupfrac}.
A fractional pupil of 1.00 would use the entire pupil and thus take advantage of the maximum JWST collecting area.
For a sub-array size encompassing only 1 DHS spectrum (such as a 2048 $\times$ 64 subarray), the largest unobscured DHS grisms should be used - element number 7 in Figure~\ref{fig:DHSvsPupilOverlay}.
Only the long-wavelength grisms of Module A are considered, because those in Module B lack an anti-reflective coating.
%We discuss only module A (of the two redundant modules A and B) because the long wavelength grisms are anti-reflective coated for module A.

@@

\subsection{Readout Considerations}\label{sec:readout}
JWST's NIRCam instrument has many different possible readout patterns per pixel and array that all factor in to the SNRs and saturation limits for an observation.
We adopt the following naming convention for the readout: an observation may consist of one of more exposures which are combined into a single \texttt{FITS} file per exposure.
Within an exposure, there are one or more integrations, which consist of a reset frame followed by one or more groups of non-destructive reads up a ramp.
In all of our examples, there is one read per group.
This is illustrated in Figure~\ref{fig:readout} for three example readout modes.

We consider a simplified example of a time series using a standard correlated double sampling (CDS) illustrated in Figure~\ref{fig:readout} and assume a round frame time of 10.0 seconds, close to the actual 10.74 seconds full frame read time.
CDS is equivalent to RAPID mode with 2 reads up the ramp.
The raw data delivered to observers for use in a pipeline would consist of a single FITS file containing a data cube.
If the data is in full frame mode, cube would have dimensions of 2048 $\times$ 2048 $\times$ 8, which can be thought of 8 image planes.
In this example there are 4 integrations consisting of a reset followed by two reads, which are saved as two image planes per integration.	
For simplicity in this example, we ignore many of the corrections that are applied in a pipeline such as non-linearity correction, reference pixel correction, intrapixel capacitance, flat fielding etc.
Under these assumptions, the flux (in DN/s) in any integration is the difference between the two reads divided by the frame time.
This reduces the 8 image planes, to 4 flux planes, one per integration.
Spectral extraction on the flux planes \citep[e.g.][]{horne1986optimalE} produces 1 flux per wavelength bin.
The time series for a given wavelength would consist of the 4 fluxes with 4 time coordinates centered at 15 sec, 45 sec, 75 sec and 105 sec if we use the mid-points of the two reads.

@@

There are several variations of readout patterns available to NIRCam with APT.
For exoplanets time series observations, the readout patterns most commonly used will likely be ``RAPID'', meaning every sample is saved up the ramp individually and ``BRIGHT1'' where a single frame is dropped between reads.
These modes give high time resolution and allow one to detect and flag cosmic ray hits without having to discard large amounts of data.
The number of groups per integration is a variable set in the APT.
When the number of groups is 1 (``1~Group Readout''), there is no ability to subtract out the bias frame so it relies on the bias level being constant through the whole time series.
When the number of groups is 2, the slope can be calculated from the difference between the frames divided by the frame time.
It is the same as the CDS mode described above when using RAPID readout.
For a higher number groups, a line is fit to the the ramp to calculate a slope image for every integration.
In exoplanet time series, the high signal to noise ratios required generally mean that bright sources will be observed so these will generally not benefit from skipped frames or co-added frames of other JWST readout modes, which come at the expense of lower time resolution, reduced ability for cosmic ray rejection and fainter saturation limits.
For more general discussion of HAWAII-2RG detector readout, we refer to \citet{rauscher2007detectors} and for specific readout patterns we refer to \citet{roberto2009optimalR1} and \citet{roberto2010optimalR2}.

@@

\begin{figure}[!h]
\includegraphics[width=1.0\columnwidth]{brightness_scape.pdf}
\caption{Saturation limits of the DHS mode (blue line), and long-wavelength grism's F322W2 filter (green line) and F444W filter (dark red line) as a function of stellar effective temperature.
The magnitudes are Vega $K_S$ magnitudes where 80\% of the well depth is filled in the standard CDS readout mode (described in Figure~\ref{fig:readout}) for a 2048 $\times$ 64 pixel sub-array using 4 amplifiers.
The blue points are known transiting systems from and the green points are simulated TESS planets \citep{sullivan2015tess}.}\label{fig:saturationLim}
\end{figure}

@@

The posterior distributions for HD 209458b in Figure~\ref{fig:corner209} show improved constraints on the temperature, cloud pressure level and the abundances of H$_2$O, CO and CO$_2$ when the long wavelength grisms are used alone (Case~\ref{it:grAlone}) as compared to when the DHS is used simultaneously with the long wavelength grisms (Case~\ref{it:DHSgr}).
These improvements come about, in part, because the 1.0$\mu$m to 2.0$\mu$m wavelength region probes to deeper pressure layers ($\sim$ 400 mbar) to exclude clouds and better constrain the scale height.
